---
title: "R Notebook"
output: html_notebook
---

Comparing means of two normally generated datasets.
```{r}
x <- rnorm(20, 2, 2)
y <- rnorm(20, 3, 2)

boxplot(x, y, horizontal = T)
```

Calculate means
```{r}
mx <- mean(x)
my <- mean(y)
```

Calculate variance
```{r}
dx <- var(x)
dy <- var(y)
```

Formula
```{r}
d <- (19*dx + 19*dy) / 38
```

```{r}
t <- (mx - my) / sqrt(d/10) 
```

t is a student distribution with degrees of freedom 38 ( this time )
Our hypothesis being mx = my
Our alternative being mx < my

Calculate p value
```{r}
pt(t, 38)
```

This was the academic way. Now with the all mighty function:
```{r}
t.test(x, y, var.equal = T, alternative = "less")
```
We can use a parameter paired of t.test as well - if the data is paired.

# Problems
### Problem 1:

Smokers yes for increasing cigarettes price:
71
All smokers - 195

Non-smokers yes for the same question:
351
All non-smokers - 605
Is both groups' opinion the same?

We use prop.test here.
Our hypothesis being: mx = my ( the opinions are the same )
Our alternative being: mx != my ( the opinions are different )
Will check if proportion 71/351 is equal to 351/605

```{r}
prop.test(c(71, 351), c(195, 605))
```
The p value is quite small here - the hypothesis cannot be true. We accept the alternative.

### Problem 2
Hypothesis: mx = my ( not better off than the ones taking placebo stuff)
Alternative: mx < my ( worse off than the placebo guys )
```{r}
new_medicament <- c(15, 10, 13, 7, 9, 8, 21, 9, 14, 8)
placebo <- c(15, 14, 12, 8, 14, 10, 7, 16, 10, 15, 12)

boxplot(new_medicament, placebo)
shapiro.test(new_medicament)

t.test(new_medicament, placebo, alternative = "less")
```
p value = 0.34 -> we accept the hypothesis

### Problem 3
```{r}
first_radar <- c(70, 85, 63, 54, 65, 80, 75, 95, 52, 55)
second_radar <- c(72, 86, 62, 55, 63, 80, 78, 90, 53, 57)

boxplot(first_radar, second_radar)
qqnorm(first_radar)
qqline(first_radar)
```
They are normally distributed. (sort of) So we will use t.test. 
Data is paired as well ( we monitor the same events twice)
Hypothesis: Both radars are equal( mfirst_radar = msecond_radar)
Alternative: Radars are different
```{r}
t.test(first_radar, second_radar, paired=T, var.equal = T)
```
P value is very high - we accept the hypothesis.

## Problem 4
```{r}
library(UsingR)

attach(ewr)

shapiro.test(AA)
shapiro.test(NW)
```
Data is not normal. So we will use wilcox test.
Hypothesis they are equal, alternative they aren't.

```{r}
wilcox.test(AA, NW)
```
P value is large enough - we can't deny the hypothesis. 

## Problem 5
https://www.itl.nist.gov/div898/handbook/eda/section3/eda3531.htm
```{r}
shapiro.test(m)
shapiro.test(n)

wilcox.test(m, n)
```
For t.test we remove all the rows with a -999 value in them and they will be normally distributed.

## Problem 6
Diodes
```{r}
first_diode <- c( 39, 50, 61, 67, 40, 40, 54)
second_diode <- c( 60, 53, 42, 41, 40, 54, 63, 69 )

boxplot(first_diode, second_diode)

shapiro.test(first_diode)
shapiro.test(second_diode)

qqnorm(first_diode)
qqline(first_diode)

t.test(first_diode, second_diode)
```
P value is big, we accept hypothesis.
Shapiro tells us they are sort of normally distributed.