---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Chi-square criterion for type of distribution

X1...XN - random variables. Hypothesis: are they of distribution F, alternative - they aren't of distribution F.

Problem 1:
H0: p1 = p2 =... = p6 = 1/6
H1: distr is not uniform

n = 150

Enter data:
```{r}
x <- c(22, 21, 22, 27, 22, 36)
```
Calculate t value ( t = sum((fi - n*pi) / n*pi) )
Where fi is the amount of values in each interval
pi - chance for the value to be in an interval
here pi is 1/6 for each i
```{r}
t <- sum((x - 25)*(x - 25) / 25)
```
This is chi-squared with 5 degrees of freedom.
We calculate the p-value.
```{r}
1 - pchisq(t, df=5)
```
P-value is big -> 0.24, so we accept the hypothesis.
With a chance of 0.24 the distribution is uniform.

There is a test for this
```{r}
chisq.test(x)
```

Problem 2

In the first 200 digits of PI, are the nine digits uniformly distributed?
```{r}
library(UsingR)
f <- pi2000[1:200]
tf <- table(f)
x <- c(19, 20, 24, 20, 22, 20, 15, 12, 25, 23)
chisq.test(x)
```
P-value is big, so we accept the hypothesis.

Problem 3

Is this text in english? 
1036 word text, x is data of occurrences of letters E,T,A,O,I,N
p is the probabilities of each of these letters occurring in an english word.
Last element of p is the sum of the prob of the other letters.

```{r}
x <- c(102, 108, 90, 95, 82, 40, 519)
p <- c(0.127, 0.0956, 0.0817, 0.0751, 0.0697, 0.0675, 0.4834)
#sum(x)
#1036 - 517

chisq.test(x, p=p)
```
p-value is small => we accept the alternative

Problem 4

Data for car crashes and consequences for people wearing seat belts and those not wearing them.

Table is of type: 
                No injury   Slight injury   Middle sized injury Heavy injury
With seat belt      fij         
Without seat belt   fij

This is chi-squared criterion for independent variables. 

```{r}
x <- matrix(c(12813, 647, 359, 42, 65963, 4000, 2642, 303), ncol = 4, nrow = 2, byrow = T)

chisq.test(x)
```
or
```{r}
z <- x[1,]
n <- x[1,] + x[2,]
prop.test(z, n)
```
p-value is small, so the variables are not independent, we accept the alternative.

Problem 5:

Quality of products by day of week: 
          M  T  W  T  F
Excellent 44 74 79 72 31
Good      14 25 27 24 10
Medium    15 20 20 23 9
Bad       3  5  5  0  0

```{r}
x <- matrix(c(44, 74, 79, 72, 31, 14, 25, 27, 24, 10, 15, 20, 20, 23, 9, 3, 5, 5, 0, 0), nrow=4, ncol=5, byrow=T)

chisq.test(x)
```
P-value is big, so we accept the hypothesis.