---
title: "Chi squared book problems"
output:
  html_document:
    df_print: paged
---

# Verzani - Simple R

## Chi-squared tests

### Problem 1: 

Check if random variables are independent with this data:

```{r}
non_block_data = c(18, 15, 5, 8, 4)
block_data = c(10, 5, 7, 18, 10)

chisq.test(data.frame(non_block_data, block_data))
```
Small p-value so we accept the alternative. The variables are dependent.

### Problem 2

Have people been in accidents over the last year.
Columns are No accident, minor accident, major accident.
Rows are age ranges.

Test of homogeneity, is there a difference in distributions based on age?

```{r}
under_18 = c(67, 10, 5)
eighteen_25 = c(42, 6 ,5)
twentysix_40 = c(75, 8, 4)
fourty_60 = c(56, 4, 6)
over_65 = c(57, 15, 1)

over_40 = fourty_60 + over_65

chisq.test(rbind(under_18, eighteen_25, twentysix_40, over_40))
```
We don't have much data, but the p value seems to be big, so we can't reject our null hypothesis that they are equally distributed.

### Problem 3

Parrotfish, grouper, tang historically have proportions of 5:3:4
A survey found data: parrotfish - 53, grouper - 22, tang - 49
Are they the same proportions as historically? 
Chi square test for goodness fit

```{r}
fish_data = c(53, 22, 49)
historic_probs = c(5, 3, 4) / 12

chisq.test(fish_data, p=historic_probs)
```
P value is 0.13, we cannot reject null hypothesis, so we accept that they still have the same proportions as historically.

### Problem 4

Admissions to UC Berkeley by gender. Is the distribution of the males admitted similar to that of the females? Test for homogeneity.

```{r}
data("UCBAdmissions")
x = ftable(UCBAdmissions) # flatten the table
```

Take admitted
```{r}
admitted = x[1:2,]

chisq.test(admitted)
```
p-value is very small, so we can accept the alternative that they are not equally distributed.

Now for rejected
```{r}
rejected = x[3:4,]

chisq.test(rejected)
```
Again a small p-value - alternative is true.