---
title: "Confidence intervals"
output: html_notebook
---

For a (1-a)100% confidence level, where for any a in (0, 1) we find a z* such that:
P (-z* < z < z*) = 1 - a ( z* is z(1-a/2))

```{r}
alpha = c(0.2, 0.1, 0.05, 0.001)
zstar = qnorm(1 - alpha/2)
zstar
```
And the reverse:
```{r}
2*(1-pnorm(zstar))
```

Finding 95% percent confidence interval that 42 people liked some brand from 100 surveyed people.

```{r}
prop.test(42, 100)
```
By default it makes a 95% conf interval.

```{r}
prop.test(42, 100, conf.level = .9)
```
t.test for normal data, wilcox.test for data that is normal ( it's a test for the median, confidence interval for the median )

# Problems

## Problem 1

```{r}
x = rnorm(15, 10, 5)

simple.z.test = function(x, sigma, conf.level=.95) {
  n = length(x)
  xbar = mean(x)
  alpha = 1 - conf.level
  zstar = qnorm(1-alpha/2)
  SE = sigma/sqrt(n)
  xbar + c(-zstar*SE, zstar*SE)
}

simple.z.test(x, 5)
``` 
Compare with t.test

```{r}
t.test(x)
```
T.test interval is always bigger than the z.test. As t distribution has a bigger mean than the normal when put side to side in a boxplot.

```{r}
x = rnorm(100); y = rt(100, 9)
boxplot(x, y)
```

## Problem 3

```{r}
wilcox.test(exec.pay, conf.int = T, conf.level = .8)
wilcox.test(exec.pay, conf.int = T, conf.level = .95)
```

## Problem 4

```{r}
qqnorm(rat)
qqline(rat)

t.test(rat, mu=110)
```
Normal data, using t.test. With the given conf interval, a 110 mean is appropriate. T.test says so, too.

```{r}
qqnorm(puerto)
qqline(puerto)

t.test(puerto, mu=270)
```
Same logic here.

## Problem 5
```{r}
qqnorm(malpract); qqline(malpract)

wilcox.test(malpract, conf.int = T, conf.level = .9)
```