---
title: "Data exercises"
output: html_notebook
---

# Exercises from Verzani R book, chapter 2

## Problem 1

Keeping track of mileage each time you fill up. Data is:

```{r}
mileage = c(65311, 65624, 65908, 66219, 66499, 66821, 67145, 67447)
x = diff(mileage)
x
```
Diff counts the number of miles between fill-ups. Let's find the min, max, mean.
```{r}
max(x)
min(x)
mean(x)
```

## Problem 2

Tracking commute times for two weeks ( 10 days )

```{r}
travel.time = c(17, 16, 20, 24, 22, 15, 21, 15, 17, 22)
max(travel.time)
mean(travel.time)
min(travel.time)
```
Apparently 24 is a mistake. Let's fix it.
```{r}
travel.time[which(travel.time == 24)] = 18
mean(travel.time)
```
How many times was commute 20 or more?
```{r}
sum(travel.time >= 20)
```
And percent of it less than 17?
```{r}
(sum(travel.time < 17) * 10) / 100
```

## Problem 3

Cell phone bills. How much was spent? And a bunch of other statistics.

```{r}
phone.bills <- c(46, 33, 39, 37, 46, 30, 48, 32, 49, 35, 30, 48)
sum(phone.bills)
min(phone.bills)
max(phone.bills)
sum(phone.bills > 40)
(sum(phone.bills > 40) * 12) / 100
```

## Problem 4

Buying a used car. Over 3 months of watching following prices:

```{r}
car.prices = c(9000, 9500, 9400, 9400, 10000, 9500, 10300, 10200)
mean(car.prices)
min(car.prices)
max(car.prices)
```

## Problem 5

Some functions in R.

```{r}
x = c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)

sum(x) / 10

log(x, base=10)

(x - 4.4) / 2.875

diff(range(x))
```
Last one is difference between min and max value. 