---
title: "Exploratory data analysis"
output: html_notebook
---

EDA is the process of looking at a data set to see what statistical inferences can be learned from that data. For univariate - is it normal, longer tailed, shorter tailed? Symmetry, skewed? Unimodal, bimodal? 

Our toolbox is:

Barplots - for categorical data
Hist - for shape of numerical distros 
Boxplots - for seeing summaries of numerical distr, comparing distributions
Normal probability plots - is data approximately normal?

```{r message=FALSE, warning=FALSE}
attach(homedata)

# hist(y1970); hist(y2000)

simple.eda(y1970); simple.eda(y2000)
```
Soo, heavily skewed, not normal. Lots of outliers. Heavily tailed.

Now, for exec.pay. 

```{r}
simple.eda(exec.pay)
```
Very heavily skewed. So let's do a log transform.

```{r}
log.exec.pay = log(exec.pay[exec.pay > 0])

simple.eda(log.exec.pay)
```

# Problems

## Problem 1

```{r message=FALSE, warning=FALSE}
attach(babies)

hist(wt, probability = T)
lines(density(wt))

qqnorm(wt)
qqline(wt)
```
Normal, sort of.

```{r}
simple.eda(gestation)
```
Skewed, kinda exponential.

```{r}
simple.eda(age)
```
Again, skewed, heavy tail.

```{r}
simple.eda(ht); simple.eda(wt1)
```
Height looks normally distributed, weight is skewed as hell.