---
title: "Mock exam"
output: html_notebook
---

## Problem 1

```{r}
library(MASS)
attach(Aids2)
```
How many people are less than 20 years old?
```{r}
sum(age < 20)
```
Gender of first five people
```{r}
sex[1:5]
```
What part of the men were infected through blood?
```{r}
men.count = sum(sex=='M')
men.blood = sum(T.categ=='blood')

men.blood / men.count
```
Graphic for death by state
```{r}
deaths.by.state = data.frame(state=state[status=='D'])
barplot(table(deaths.by.state), beside = T, legend=c(levels(state)), main='deaths by state')
```

## Problem 2
Is it true that women death rate is lower? Let's check a hypothesis.

```{r}
women.deaths = c(length(status[status=='D' & sex == 'F']), sum(sex == 'F'))
men.deaths = c(length(status[status=='D' & sex == 'M']), sum(sex == 'M'))

prop.test(women.deaths, men.deaths, alternative = 'less')
```
Big p-value. We can't decline the hypothesis. Women death rate is not lower.

## Problem 3
Is the mean 38?

```{r}
dead.age = age[status=='D']

qqnorm(dead.age)
qqline(dead.age)

shapiro.test(dead.age)
```
Definitely not normal. So wilcox?

```{r}
wilcox.test(dead.age, mu = 38)
```
Well, it's not 38, apparently. ( There are a lot of outliers )

## Problem 4
 
```{r}
x.sample = rchisq(100, df=10)
hist(x.sample, probability = T)

lines(sort(x.sample), dchisq(sort(x.sample), df=10))
```

## Problem 5
Relation between body weight and heart weight for cats.

```{r, message=FALSE, warning=FALSE}
attach(cats)

male.cats = cats[cats$Sex == 'M',]
Bwt = male.cats$Bwt
Hwt = male.cats$Hwt

lm.heart = lm(Hwt ~ Bwt)
```
So the relation is Hwt = 4.31 * Bwt - 1.18

```{r}
new.weights = male.cats$Bwt + 1
new.heart.weights = predict(lm.heart, data.frame(Bwt=new.weights))

lm2.heart = lm(new.heart.weights ~ Bwt)

lm2.heart$coefficients
```
Heart is heavier by about 4 gr when weight is more. So it's not 5.

Confidence interval for Hwt for Bwt = 2.6 kg.
```{r}
predict(lm.heart, data.frame(Bwt=2.6), level = .95, interval = "confidence")[2:3]
```

## Problem 6

Three groups of 20 women, three different diets. Comparing the data of how much weight they lost.

```{r}
diet.data = read.csv("./Data1.txt", sep=";", dec=",")

hist(diet.data, probability = T)

shapiro.test(diet.data$x1)
shapiro.test(diet.data$x2)
shapiro.test(diet.data$x3)
```
We are going to compare them two by two. For all three columns, we don't reject null hypothesis that they are normally distributed.

```{r}
t.test(diet.data$x1, diet.data$x2)
t.test(diet.data$x2, diet.data$x3)
t.test(diet.data$x1, diet.data$x3)
```

This can be done with anova.

```{r}
diet.vector = c(diet.data$x1, diet.data$x2, diet.data$x3)
n = rep(20, 3)

group = rep(1:3, n)

diet.data2 = data.frame(data=diet.vector, group=factor(group))

fit = lm(diet.vector ~ group, diet.data2)

anova(fit)
```
Not really sure how correct this is, but worth a shot.