---
title: "Multivariate data"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
attach(PlantGrowth)

# unstack makes a new data frame by the levels of the factor
boxplot(unstack(PlantGrowth))

# same thing, a.k.a weight by the variable group
boxplot(weight ~ group)
```

# Problems

## Problem 1
```{r, message=FALSE, warning=FALSE}
attach(emissions)

plot(GDP, CO2)
# identify(GDP, CO2, n = 1)
# identified 1 as outlier

plot(GDP[-1], CO2[-1])
```

## Problem 2

Make a side-by-side boxplot of the chips dataset.
```{r message=FALSE, warning=FALSE}
attach(chips)

boxplot(chips)
```
The means look pretty similar. 

## Problem 3

Same thing for chickens.

```{r message=FALSE, warning=FALSE}
attach(chicken)

boxplot(chicken)
```
So, we can see that with the third ration the chickens are the fattest. In fact, we see a steady rise of the means with 1 kg with each ration.

## Problem 4
```{r message=FALSE, warning=FALSE}
attach(kid.weights)

# age.yr = cut(age, seq(0, 144, by=12), labels=0:11)

# age = age.yr

boxplot(weight ~ age)
```
So, there is a steady rise up until the 10th year. Then children tend to lose weight.

## Problem 5
```{r message=FALSE, warning=FALSE}
attach(carbon)

# side by side boxplot of monoxide reading by site
boxplot(unstack(carbon))
```
Levels at second site are quite larger.

## Problem 6
```{r, message=FALSE, warning=FALSE}
attach(babies)

# pairs(babies)

# weight and gestation depending on the level of smoking
plot(wt, gestation, pch=smoke)
```
So many variables...