---
title: "Two Sample tests"
output: html_notebook
---

Verzani, Simple R, chapter 11.
Two surveys. 

Week 1 -> Favorable 45, Unfavorable 35
Week 2 -> 56, 47

Are they equal?

```{r}
prop.test(c(45, 56), c(45+35, 56+47))

# or

chisq.test(matrix(c(45, 56, 35, 47), ncol=2, nrow=2))
```
Well, not much of a difference then, in the two weeks.

Two sample t tests

t statistic now is (x1 - x2) - (mu1 - mu2) / sqrt(s1^2 / n1 + s2^2 / n2), where x1 and x2 are averages.

When we assume equal variances...
```{r}
with.drug = c(15, 10, 13, 7, 9, 8, 21, 9, 14, 8)
placebo = c(15, 14, 12, 8, 14, 7, 16, 10, 15, 12)
```
So we try with equal variances...
```{r}
t.test(with.drug, placebo, alt="less", var.equal = TRUE)
```
It's not very different with assumption of unequal variances...
```{r}
t.test(with.drug, placebo, alt="less")
```

For paired samples: We assume the variables are dependent on one another. We assume that Y is X + e ( some vector of errors ). We test the mean of the errors: whether it is 0, as in, the data is similar. 

```{r}
grader.one = c(3, 0, 5, 2, 5, 5, 5, 4, 4, 5)
grader.two = c(2, 1, 4, 1, 4, 3, 3, 2, 3, 5)

t.test(grader.one, grader.two, paired=T)
```
Reject the null hypothesis.

# Problems

## Problem 1

```{r}
attach(homework)

boxplot(Private, Public)

t.test(Private, Public, paired=T)

detach(homework)
```
We can't reject hypothesis.

## Problem 2

```{r message=FALSE, warning=FALSE}
attach(corn)

# without matched sample

t.test(Standard, New)

# with matched sample

t.test(Standard, New, paired=T)
```
Reject the null hypothesis.

## Problem 3

```{r}
attach(blood)

t.test(Machine, Expert)
```
They seem to be equal.