---
title: "Univariate data"
output: html_notebook
---

# Exercises from chapter 3 of the Verzani R book

First, dabblings while I read the chapter:
Categorical data.
Smoking survey, for example.

```{r}
x = c("Yes", "No", "No", "Yes", "Yes")
table(x)
```
So it adds up the frequency of each element. Nice. Now factors. They are a way to classify categorical data into levels.
```{r}
factor(x)
```
Bar charts. Shows frequencies and proportions. Types of beers that people drink.
```{r}
beer = c(3, 4, 1, 1, 3, 4, 3, 3, 1, 3, 2, 1, 2, 1, 2, 3, 2, 3, 1, 1, 1, 1, 4, 3, 1)
barplot(beer)
barplot(table(beer))
barplot(table(beer)/length(beer))
```
What else?
```{r}
pie(table(beer), col=c("purple", "green2", "cyan", "white"))
```
Numerical data - summary. Mean, var, sd, median, fivenum, summary. Quantiles, quartiles.
p quantile is 1 + (n-1)p. The .25, .5, .75 quantiles are called quartiles.
```{r}
fivenum(beer)
```

```{r}
data = c(10, 17, 18, 25, 28, 28)

summary(data)
quantile(data, c(.25, .5, .75))
```
Remove some outliers from data.

```{r}
data[7] = 284728481748

mean(data, trim=1/10)
mean(data, trim=2/10)
median(data)
mean(data, trim=1/2)
```
Measure of spread. ( as in var and sd for normal distr ) mad() and IQR()
Find values in intervals - cut(breaks=c(intervals))

```{r}
hist(data, probability = T)
rug(jitter(data)) # shows points
# lines(density(data))
simple.freqpoly(data)
```
Boxplots!

```{r}
boxplot(data, horizontal = T)
```

```{r, warning=FALSE, message=FALSE}
data("faithful")
attach(faithful)

hist(eruptions, 15, prob=T)
lines(density(eruptions))
```

# Problems

## Problem 3
The first two are strange.

Generating samples with rnorm.

```{r}
x = rnorm(100)
y = rnorm(100)

hist(x, probability = T)
lines(density(x))

hist(y, probability = T)
lines(density(y))
```

## Problem 4

```{r}
library(UsingR)

hist(south, probability = T)
boxplot(south, horizontal = T)
```
This one is skewed to the left, has some outliers.

```{r}
hist(crime$y1983, probability = T)
boxplot(crime$y1983, horizontal = T)
```
Quite the same here.

```{r}
hist(aid, probability = T)
boxplot(aid, horizontal = T)
```
This one is more symmetric and not skewed.

## Problem 5
Just looking at data sets. 
```{r}
hist(math, probability = T)
hist(firstchi, probability = T)
hist(bumpers, probability = T)
```

## Problem 6
O-ring failures for first 23 flights of us space shuttle Challenger

```{r}
shuttle.failures = c(0, 1, 0, NA, 0, 0, 0, 0, 0, 1, 1, 1, 0, 0, 3, 0, 0, 0, 0, 0, 2, 0, 1)

table(shuttle.failures)

mean(shuttle.failures, na.rm=T)
```

## Problem 7 
Yay, pie.

```{r}
hist(pi2000, breaks=c(0:9), probability = T)

# proportion of 1s
sum(pi2000 == 1) / 2000

# of 2s
sum(pi2000 == 2) / 2000

# of all
table(pi2000) / 2000

lines(density(pi2000))
```