---
title: "R Notebook"
output: html_notebook
---

Linear regression

```{r}
x = runif(20, 1, 10)
y = 2*x + 3 + rnorm(20, 0, 2)
plot(x, y)
```
Linear model for data
```{r}
# y is a function of x
l = lm(y ~ x)

# The plots are 4, we can show the one we want with second arg of plot.
plot(l, 1)

# How much a point is far from the line
l$residuals

l$fitted.values
```

Now a bad model
```{r}
z = x * x + 1 + rnorm(20, 0, 2)
plot(x, z)

l2 = lm(z ~ x)
abline(l2)

# So not a good model, too big residuals, line is not horizontal
plot(l2, 1)
```

Our favourite qqplot
```{r}
plot(l2, 2)

plot(l, 2)
```

Now with data that is not a function of x
```{r}
w = runif(20, 1, 10)
plot(x, w)

l3 = lm(w ~ x)

abline(l3)

plot(l3, 1)
plot(l3, 2)
```

Summary of linear model
```{r}
# Of good model
summary(l)

# Of bad model
summary(l3)
```

Checking if beta 0 is equal to 2 in the first model
H0: B1 = 2
H1: B1 != 2

```{r}
sm = summary(l)
t = (l$coefficients[2] - 2) / sm$coefficients['x', 'Std. Error']

# calculate p value, 20 - 2 df
pt(t, 18, lower.tail = F) * 2
```
Big p value, we can't decline hypothesis.

```{r}
sm = summary(l)
sm$r.squared
```

## Problems

#### Problem 1
Y - Pulse
X - Age

Hypothesis Y = 220 - X

```{r}
x = c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
y = c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 183, 178)

model = lm(y ~ x)
plot(model, 1)

sm = summary(model)
```
Model looks like linear, both p-values are small, so none of the coefficients are zero.

We check the hypothesis Y = 220 - X
H0 : B1 = -1
H1 : B1 != 1

```{r}
standard.error = sm$coefficients['x', 'Std. Error']

# Calculate the statistic 
# model$coefficients are the coeffs of the regression
# sm$coefficients are the summary coefficients
t = (model$coefficients[2] + 1) / standard.error
t

# Calculate p-value
p.value = pt(t, 13, lower.tail = F) * 2
p.value
```

Now predicting pulse for 30, 40, 50 years
```{r}
d = data.frame(x = c(30, 40, 50))

predict(model, d)
```

90% Confidence interval for the pulse
```{r}
pr = predict(model, d, interval = 'confidence', level = 0.9)
plot(x, y)

lines(d$x, pr[,2], col = 'red')
```

Prediction for the pulse
```{r}
df = data.frame(x = seq.int(min(x), max(x), 5))
pr = predict(model, df, interval = 'confidence', level = 0.9)
pr

plot(x, y)
lines(df$x, pr[,3], col='red')
```

#### Problem 2

```{r}
data(mtcars)
```